
\chapter*{Acknowledgments}

There once was a time I thought it was some happy accident that I kept finding myself in exactly the right place.
Be it college, career, or concert, it always felt like gravity, quietly inevitable, pulling me into the things I should do, the places I should go, the opportunities I should pursue.
But now, on the backside of this arc, I can see quite plainly that I always had the direction wrong; it was never a pull, but a \emph{push}.
Come to think of it, for as long as I can seem to recall, I have been lucky enough to be surrounded by those who would inspire, encourage, force, drag, and challenge me to always be better.
I am nothing short of blessed to be graced by the company and influence of such wonderful human beings, and will try my best to thank as many of you as I can here.

% Home
First and foremost, I am forever grateful to my grandparents ---Glendolyn \& Norman, Elizabeth \& Frank--- for their myriad contributions, both concrete and intangible.
It is only for their love, diligence and sacrifice that any of this is possible (but what I wouldn't give to see your reactions now).
This goes doubly so for my parents, Sharon \& Branden, who took a computer-tinkering, dinosaur-building, saxophone-wielding dreamer and turned him into a computer-sciencing, wood-working, guitar-wielding dreamer.
Good job, guys.
An extra special thanks goes to my brothers, Steve \& Tim, for all adventures, past, present, and future, as we continue down the strangest of trails.

% Rhody / Cuse
To those who had a most profound impact on my younger self:
Paul Andersen, who helped introduce me to music, and more importantly, improvisation;
Mary Bolton, who truly was ``Mary E. Best'', and is near entirely to thank (or blame) for my egregious use of analogy;
Matt McGuire, who instilled the virtues of hard work, effort, and discipline, and earned my deepest respect for it;
and Jayant Datta, who saw something in a kid with long hair and an unhealthy love of delay pedals, and happily pushed him down the rabbit hole of higher education.

% Miami
To my professors and colleagues from my time in Miami for helping me grow into a proper MuE.
I cannot thank Corey Cheng and Colby Leider enough, the former for his focused direction and the latter for the opposite.
To my classmates, you have influenced more than you know, both then and now:
without the positive influence of Chris Santoro I would be a different person I am today;
Estefania Cano, for setting the highest bars of scientific prowess, musicality, and humility;
Glen Deslauriers, who brought the creativity, heart, and synths in spades;
and to Patrick O'Keefe and Reid Draper, who helped stoke the coals of ambition into a blaze.

% NYU
To my friends and mentors ---and those that are both--- from NYU:
thanks to Aron Glennon, for conversations that always pushed the boundaries of what either of us understood;
to Areti Andreopoulou, Taemin Cho, Jon Forsyth, whose experience and knowledge helped guide me through the deep, dark forest of doctoral studies.
thanks to Oriol Nieto and Braxton Boren, without whom I cannot imagine getting this far, nor do I want to;
to Finn Upham, Michael Musick, Rachel Bittner, and Andrew Telichan, for your diverse perspectives, curious minds, and consistently sunny dispositions;
and, to Brian McFee and Justin Salamon, who even now continue to push me out into deeper waters to swim alongside them.

A special thanks to Ron Weiss, Ryan Rifkin, and Dick Lyon for the opportunity of a lifetime;
working with you was as enjoyable as it was formative, and I am forever in your debt.

Finally, a special thank you to my committee:
to Panayotis Mavromatis, whose expertise across various facets of music has been instrumental in the course of my studies;
to Yann LeCun, for sharing your wisdom and knowledge, and showing me the ways of the Force;
and to Juan Pablo Bello, for taking me under his wing, being the perfect counterweight, and showing me how to devine clarity from chaos.
You have been a fantastic advisor and friend, and I can only hope the things I've learned will generalize in the wild.


